test

### 2014-12-17 talks/test.Rmd
added some talks

---
title: "Test-Presentation"
author: "Matthew Stephens"
date: "September 23, 2014"
output: ioslides_presentation
---

% False Discovery Rates, A New Deal
% Matthew Stephens
% 2014/7/8

```{r, include=FALSE}
require("qvalue")
require("ashr")
require("ggplot2")
```

```{r setup, include=FALSE}
  #set global chunk options
  #opts_chunk$set(cache=TRUE,autodep=TRUE,warning=FALSE)
  #dep_auto()
```

# Before we get started: Let's get Organized!

- Over ~10 years of working with graduate students + postdocs,
I've noticed something.
- Organized researchers get more done (and better!).
- Many of them are more organized than I am!
- Thought: I should get organized; I should help others get organized.

# So what can you do?

- Buy a notebook; bring it to meetings; make notes! 
- Come to meetings with a written agenda. 
- While doing research, record what you did and what the outcome was.
- Use version control ([git](http://git-scm.com)) and internet repositories ([bitbucket](http://www.bitbucket.org), [github](http://www.github.com)) to organize notes, code, etc.
- Use *knitr* to help make your research reproducible.
- Talk about the tools you find useful!

# What are these repository things?

- A repository: a central place in which an aggregation of data is kept and maintained in an organized way (searcharticle.com)
- Great for sharing material across multiple people (eg student and advisor!)
- An amateur example: [http://github.com/stephens999/ash](http://github.com/stephens999/ash)

# What is knitr?

- An R package
- A tool for literate programming
- Text, and R code are interleaved
- When you compile the document, the code is run, and output inserted into the text.
- Great for writing reports, and keeping a track of what you did and what the result was!
- This talk was written with knitr (with RStudio)!

# What is Reproducible Research?

- Principle: when publishing results of computational procedures, we should
publish the code that produced the results.
- "publishing figures or results without the complete software environment could
be compared to a mathematician publishing an announcement of a mathematical theorem without giving the proof" (Buckheit and Donohoe)
- “an article about a computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result.” [Claerbout]

# Why is reproducibility important?

- Not only because people are forgetful, error-prone, or dishonest!
- Reproducing work is also the first step to extending it.
- Helps communications among researchers (eg student + advisor).
- If you do not publish code implementing your methods, your methods will likely go unused.

# More on git, github, knitr, reproducibility

- Google "The git book", to get started on git.
- Google "Karl Broman github tutorial" for statistics-oriented intro to github.
- Google "donohoe buckheit" for "Wavelab and reproducible research"

# The Canonical Genomics Experiment

- Measure lots of things, with error

- Get estimates of effects $\beta_j$ ($\hat\beta_j$) and their standard errors $s_j$

- Turn these into Z-scores, $z_j = \hat\beta_j/s_j$

- Turn these into $p$ values, $p_j$

- Apply `qvalue` 
to identify findings ``significant" at a given FDR.

- ...?

# FDR, local fdr, and q values

Although precise definitions vary depending on whether one
takes a Bayesian or Frequentist approach to the problem, roughly

- The FDR at a threshold $P$ is 
$$\text{FDR}(P)=\Pr(\beta_j = 0 |  p_j<P).$$

- The q value for observation $j$ is $q_j=\text{FDR}(p_j)$.

- The local false discovery rate, fdr, at threshold $P$ is 
$$\text{fdr}(P) = \Pr(\beta_j =0 | p_j=P ).$$

- The fdr is more relevant, but slightly harder to estimate than 
FDR because it involves density estimation rather than tail-area estimation.

```{r, echo=FALSE}
  #simple simulated example
  ncz = 100 # number of bins in z score histograms
  set.seed(111)
  hh.betahat = rnorm(1000,0,2)
  hh.sebetahat = 1
  hh.zscore = hh.betahat/hh.sebetahat
  hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)
  hh.ash = ash(hh.betahat,hh.sebetahat, method="fdr")
  hh.q = qvalue(hh.pval)
```

```{r, echo=FALSE}
  plot_FDReg_hist = function(hh.pval,pi0,nc=40,nullcol="blue",altcol="cyan",type=4,title="Distribution of p values",...){
    hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main=title,nclass=nc,col=altcol,...)
    if(type>1){
      abline(h=pi0,col=nullcol,lwd=2)

      hh.hist$density=rep(pi0,length(hh.hist$density))  
      #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
      plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)
    }
    if(type>2){
    abline(v=0.1,lwd=2,col=2)
    }
    if(type>3){
      text(0.05,1.2,labels="A",col=2,cex=1.2)  
      text(0.05,0.4,labels="B",col=2,cex=1.2)  
      text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
    }
  }
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=1)                 
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=2)                 
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=3)                 
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4)                 
```

# Is this an important problem?

- The original paper introducing FDR (Benjamini and Hochberg, 1995) has
been cited 21,787 times (May 2014) according to Google Scholar.

- That is three times a day for the last 19 years!

# Problem 1: The Zero Assumption (ZA)

- The standard `qvalue` 
approach assumes that all the $p$ values near 1 are null.

- Analogously, one can assume that all Z scores near 0 are null. Efron refers to this as the ``Zero Assumption".

- Seems initially natural.

```{r, echo=FALSE, include=FALSE}
require(fdrtool)
  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
require(locfdr)
  hh.locfdr = locfdr(hh.zscore,nulltype=0,plot=0)
require(mixfdr)
  hh.mixfdr = mixFdr(hh.zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
```

# Implied distribution of $p$ values under $H_1$

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4)                 
```

# Implied distribution of Z scores under alternative

```{r, echo=FALSE,fig.height=4,fig.cap=""}
#plot a histogram of z scores, highlighting the alternative distribution
#of z scores that is implied by localfdr values lfdr.
  nullalthist = function(z,lfdr,nullcol="blue",altcol="cyan",...){
    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
    h$density = (1-avlfdr) * h$density
    plot(h,add=TRUE,col=altcol,freq=FALSE)
  }

  nullalthist(hh.zscore,hh.fdrtool$lfdr)  
```

# FDR problem 2: different measurement precision

- In some cases the measurement precisions differ among units

- Eg Expression levels of low-expressed genes have less precision than high-expressed genes

- If some effects are measured less precisely than others, those tests ``lack power"
and dilute signal, increasing FDR

# Example: Mouse Heart Data

```{r, echo=FALSE}
x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
xx = rowSums(x[,2:5])
x = x[xx>0,]
xx = xx[xx>0]
```

```{r echo=FALSE}
head(x)
```

- Data on 150 mouse hearts, dissected into left and right ventricle
(courtesy Scott Schmemo, Marcelo Nobrega)

```{r, echo=FALSE,include=FALSE}

load("../scratchwork/Smemo_qbinom_zdat.RData")
# two-sided test
  ttest.pval = function(t, df) {
    pval = pt(t, df = df, lower.tail = T)
    pval = 2 * ifelse(pval < 0.5, pval, 1 - pval)
    return(pval)
}

tscore = zdat[3, ]/zdat[4, ]
pval = ttest.pval(tscore, df = 2)
qval = qvalue(pval)

highxp = xx>1000 # select high expressed genes
pval.high = pval[highxp]
qval.high = qvalue(pval.high)

zdat.ash = ash(zdat[3,],zdat[4,],df=2,method="fdr",mixcompdist="halfuniform")
zdat.ash.high = ash(zdat[3,highxp],zdat[4,highxp],df=2,method="fdr",mixcompdist="halfuniform")

```

# Example: Mouse Heart Data

```{r, echo=FALSE,fig.height=4,fig.cap=""}
qval= qvalue(pval)
plot_FDReg_hist(pval,pi0=qval$pi0,ylim=c(0,4))
```

# Mouse Data: Counts vary considerably across genes

```{r, echo=FALSE,fig.height=4,fig.cap=""}
hist(log10(xx),main="Distribution of total counts", xlab="log10(counts)")
```

# Lower count genes, less power

```{r, echo=FALSE,fig.height=4,fig.cap=""}
qval.low=qvalue(pval[!highxp])
pval.low=pval[!highxp]
plot_FDReg_hist(pval.low,pi0=qval.low$pi0,ylim=c(0,4))
```

# Higher count genes, more power
```{r, echo=FALSE, fig.height=4,fig.cap=""}
plot_FDReg_hist(pval.high,pi0=qval.high$pi0,ylim=c(0,4))
```

# Low-count genes dilute signal at high-count genes
```{r, echo=FALSE, fig.height=4,fig.cap=""}
plot_FDReg_hist(pval,pi0=qval$pi0,ylim=c(0,4))
```

# FDR problem 2: low count genes add noise, increase q values
```{r,echo=FALSE, fig.height=4,fig.cap=""}
plot(qval.high$qval,qval$qval[highxp],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
abline(a=0,b=1,col=2)
```

# Problems: Summary

Standard tools are unduly conservative.

- The ZA, which implies actual effects have a (probably unrealistic) bimodal distribution; causes overestimate of $\pi_0$, losing power.

- By focussing on $p$ values, low-precision measurements can dilute high-precision measurements.

# FDR via Empirical Bayes

- Following previous work (e.g. Newton, Efron, Muralidharan) we take an empirical Bayes approach to FDR.

- Eg Efron assumes that the $Z$ scores come from a mixture of null, and alternative:
$$Z_j \sim f_Z(.) = \pi_0 N(.;0,1) + (1-\pi_0) f_1(.)$$
where $f_1, \pi_0$ are to be estimated from the data.

- Various semi-parametric approaches taken to estimating $f_1$. For example,
Efron uses Poisson regression; Muralidharan uses mixture of normal distributions.

- Once $f_1$ and $\pi_0$ estimated, FDR calculations are straightforward.

# FDR: A New Deal

- Instead of modelling $Z$ scores, model the effects $\beta$,
$$\beta_j \sim \pi_0 \delta_0(.) + (1-\pi_0) g(.)$$

- Constrain $g$ to be unimodal about 0; estimate $g$ from data.

- *Incorporate precision* of each observation $\hat\beta$ into the likelihood.
Specifically, approximate likelihood for $\beta_j$ by a normal: 
$$L(\beta_j) \propto \exp(-0.5 (\beta_j - \hat\beta_j)^2/s_j^2).$$
[From $\hat\beta_j \sim N(\beta_j, s_j)$]
Or, better, use a $t$ likelihood if $s_j$ estimated using few observations.

# FDR - A New Deal

- A convenient way to model $g$: mixture of 0-centered
normal distributions: 
$$g(\beta; \pi) = \sum_{k=1}^K \pi_k N(\beta; 0, \sigma^2_k)$$

- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (EM algorithm).

- By allowing $K$ large, and $\sigma_k$ to span a dense grid of values,
we get a flexible unimodal symmetric distribution.

- Can approximate, arbitrarily closely, any scale mixture of normals.
Includes almost all priors used for sparse regression problems (spike-and-slab, double exponential/Laplace/Bayesian Lasso, horseshoe).

# FDR - A New Deal

- Alternatively, a mixture of uniforms, with 0 as one end-point of the range,
provides still more flexibility, and in particular allows for asymmetry.

- If allow a very large number of uniforms this provides the non-parametric mle for $g$; cf Grenander 1953; Cordy + Thomas 1997.

# Illustration: $g$ a mixture of 0-centered normals

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, dnorm(x,0,1),type="l",ylim=c(0,2),ylab="density")
lines(x, dnorm(x,0,0.1))
lines(x, dnorm(x,0,0.5))
```

# Illustration: $g$ a mixture of 0-centered normals

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, 0.5*dnorm(x,0,1)+0.5*dnorm(x,0,0.1),type="l",ylim=c(0,2),ylab="density")
```

# Illustration: $g$ a mixture of 0-anchored uniforms

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, dunif(x,0,1),type="s",ylim=c(0,5),ylab="density")
lines(x, dunif(x,0,0.2),type="s")
lines(x, dunif(x,0,0.5),type="s")
lines(x, dunif(x,-0.3,0),type="s",col=2)
lines(x, dunif(x,-0.4,0),type="s",col=2)
lines(x, dunif(x,-2,0),type="s",col=2)
```

# Illustration: $g$ a mixture of 0-anchored uniforms

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, 0.1*dunif(x,0,3)+ 0.3*dunif(x,0,0.2) + 0.2*dunif(x,0,0.5) + 0.1* dunif(x,-0.3,0) + 0.1*dunif(x,-0.3,0)+0.2*dunif(x,-0.4,0),type="s",ylim=c(0,2),ylab="density")
```

# Adaptive Shrinkage

- This approach actually provides a full posterior distribution for each $\beta_j$.

- So easy to obtain point estimates and credible intervals.

- Because $g(\beta)$ is unimodal, the point estimates (and CIs) will tend to be ``shrunk" towards the overall mean (0).

- Because $g(\beta)$ is estimated from the data, the amount
of shrinkage is adaptive to signal in the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each observation.

- So we call the approach ``Adaptive Shrinkage" (ASH).

# Recall Problem 1: distribution of alternative Z values multimodal
```{r, echo=FALSE,fig.height=4,fig.cap=""}
nullalthist(hh.zscore,hh.fdrtool$lfdr,main="qvalue")  
```

# Problem Fixed: distribution of alternative Z values unimodal
```{r, echo=FALSE,fig.height=4,fig.cap=""}
nullalthist(hh.zscore,hh.ash$lfdr,main="ash")
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4,title="qvalue")                 
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,get_pi0(hh.ash),type=4,title="ash")                 
```

# Recall Problem 2: low count genes add noise, increase q values
```{r,echo=FALSE, fig.height=4,fig.cap=""}
plot(qval.high$qval,qval$qval[highxp],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
abline(a=0,b=1,col=2)
```

# Problem Fixed: incorporating precision reduces influence of low-count genes
```{r,echo=FALSE, fig.height=4,fig.cap=""}
plot(zdat.ash.high$qvalue,zdat.ash$qvalue[highxp],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
abline(a=0,b=1,col=2)
```

# A new problem: an embarrassment of riches

- If the null is mostly false, the new approach can  provide unsettling results

- The FDR can be small for all observations, even those with $p \approx 1$!

- In the illustrative example, the maximum $q$ value is `r round(max(hh.ash$qvalue),2)`

# Perhaps we didn't really understand the question?

- Problem arises only if we insist on asking question ``is $\beta_j=0$?"

- Given enough signal, we become convinced that very few of the $\beta_j=0$

- But for some $\beta_j$ we still may have little information about actual value

- Suggests a change of focus: ask for which $\beta_j$ are we confident about the sign (cf Gelman et al, 2012).

# The False Sign Rate

- Suggestion: replace FDR with local false sign rate (lfsr), the probability that if we say an effect is positive (negative), it is not.

- Example: suppose we estimate that $\Pr(\beta_j<0)=0.95, \Pr(\beta_j=0) = 0.025$ and $\Pr(\beta_j>0)=0.025$. Then we report $\beta_j$ as a ``(negative) discovery", and estimate its lfsr as 0.05.

# Even with many signals, large $p$ values have high lfsr

```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
  #res=data.frame(p=hh.pval, lfsr=hh.ash$lfsr, qvalue = hh.ash$qvalue)
  library("reshape2")
  res=data.frame(p=pval, lfsr=zdat.ash$lfsr, qvalue = zdat.ash$qvalue)
  res.melt = melt(res, id.vars=c("p"),variable.name="Measure")
  cbbPalette <- c("#000000", "#D55E00", "#CC79A7")
  labels = c('qvalue','lfsr')
  breaks = c("qvalue","lfsr")

  pp= ggplot(data=res.melt,aes(p,value,color=Measure)) +geom_point(shape=1) +
      geom_abline(colour = "black") +
        xlab("p-value") +
        ylab("lfsr/qvalue")

  print(pp +scale_y_continuous(limits=c(0,1)) +
          scale_x_continuous(limits=c(0,1))  +
           scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
          coord_equal(ratio=1))

```

# Advantages of False Sign Rate

- The FDR is typically not actually identifiable from data.

- This is because data cannot distinguish between $\beta_j = 0$ and $\beta_j$ "very
small". So $\pi_0$ is not identifiable, and FDR is very sensitive to $\pi_0$.

- So methods for estimating $\pi_0$ and FDR, including those presented here, are designed to be ``conservative" (i.e. overestimate the FDR).

- The False Sign Rate is much less senstive to $\pi_0$, and hence more identifiable from data!

# Simulated example: $\pi_0$ not identifiable.

```{r,echo=FALSE,include=FALSE}
load("../paper/Rcode/sim1.RData") #load simulation results
source("../paper/Rcode/plot_pi0.R")
```

```{r, echo=FALSE,fig.height=4,fig.cap=""}
plot_pi0(list(simres1a))
```

# Simulated Example: so fdr not identifiable

```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
source("../paper/Rcode/plot_lfsr.R")
  plot_lfsr(list(simres1a),0.1,ptype="lfdr")
```

# Simulated Example: fsr much more identifiable

```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
  plot_lfsr(list(simres1a),0.1,ptype="lfsr")
```

# Summary

- ASH provides a generic approach to shrinkage estimation, as well as
false discovery (sign) rates.

- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values or $z$ scores) varying precision of 
measurements is better accounted for.

- Unimodal assumption for effects reduces conservatism

- False Sign Rate preferable to False Discovery Rate: more identifiable, and
better representation of information in data for ``high-signal" situations.

# Other Applications

- Widely applicable: requiring only an estimated
effect size and standard error for each object.

- E.g. Currently applying it to wavelet shrinkage applications.

# Next steps?

- Incorporate shrinkage of variances and not just means. (e.g. ``moderated $t$ test", Smyth et al)

- Allow $g(\cdot;\pi)$ to depend on covariates $X$.

- Allow for correlations in the measured $\hat\beta_j$.

# Thanks

- to the developers of **R**, **knitr**, **Rstudio** and **Pandoc**.

- to the several postdoctoral researchers and students
who have worked with me on related topics.

- Including Ester Pantaleo, Scott Powers, Mengyin Lu, Sen Tian, Wei Wang, Zhengrong Xing.

- NHGRI for funding.

- `ashr` package: `http://www.github.com/stephens999/ash`

# Pandoc Command used

`pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  slides.md -o slides.pdf`

(alternative to produce html slides; but figures would need reworking)
`pandoc -s -S -i -t dzslides --mathjax slides.md -o slides.html`

Here is my session info:

```{r session-info}
print(sessionInfo(), locale=FALSE)
```

# Some odd things in the data

```{r, echo=FALSE,fig.height=3}
  plot(zdat[3,],zdat[4,],ylab="standard error",xlab="beta-hat")
  dd[tail(order(zdat[3,])),]
```

# A technicality

- Suppose you estimate $\Pr(\beta_j<0)=0.98$,  $\Pr(\beta_j>0)=0.01$, $\Pr(\beta_j=0) = 0.01$. 
- Should you declare an fdr of 0.01 or 0.02?
- Maybe fsr makes more sense anyway?

# Adaptive Shrinkage of point estimates

- Recall idea: amount of shrinkage depends on measurement precision, $s_j$.

# Adaptive Shrinkage of point estimates

```{r, fig.height=4,echo=FALSE,fig.cap=""}
  res=data.frame(betahat =zdat[3,], Estimate=zdat.ash$PosteriorMean,highxp=highxp)
  cbbPalette <- c("#000000", "#D55E00", "#CC79A7")
  labels = c('betahat','Estimate')
  breaks = c("betahat","Estimate")

  pp= ggplot(data=res,aes(betahat,Estimate,col=as.factor(highxp))) +geom_point(shape=1) +
      geom_abline(colour = "black") +
        xlab("betahat") +
        ylab("shrunk betahat")

  print(pp +scale_y_continuous(limits=c(-2,2)) +
          scale_x_continuous(limits=c(-2,2))  +
           scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
          coord_equal(ratio=1))

#plot(zdat[3,],zdat.ash$PosteriorMean,main="Less shrinkage for high-count genes (red)",xlab="betahat",ylab="shrunk betahat",xlim=c(-2,2),ylim=c(-2,2), col=1+highxp)
#abline(a=0,b=1,col=2)
```

# Shrinkage is adaptive to information

Need to fix counts.associate to use fdr method in ash

```{r, echo=FALSE,fig.height=4}
#hist(zdat[3,])
#plot(zdat[3,],zdat.ash$PosteriorMean,xlim=c(-0.2,0.2))
#points(zdat[3,16677],zdat.ash$PosteriorMean[16677],col=2,pch=16)
#points(zdat[3,16079],zdat.ash$PosteriorMean[16079],col=2,pch=16)
#x[16677,]
#x[16079,]
```

```{r, echo=FALSE}
#temp = counts.associate(cc,c(-1,1,-1,1),1)
#temp.tscore = temp$zdat[3,]/temp$zdat[4,]
#temp.pval = ttest.pval(temp.tscore,df=2)
#plot(temp$zdat.ash$localfdr,temp.pval)
#identify(temp$zdat.ash$localfdr,temp.pval)
```

```{r, echo=FALSE,fig.height=3,fig.cap=""}
plot(zdat.ash$lfdr,pval,ylab="p value", xlab="ASH local fdr")
```

# Shrinkage is adaptive to information

```{r, echo=FALSE,fig.height=3,fig.cap=""}
plot(zdat.ash$lfdr,pval,ylab="p value", xlab="ASH local fdr")
points(zdat.ash$lfdr[15325],pval[15325],col=2,pch=16)
points(zdat.ash$lfdr[16123],pval[16123],col=2,pch=16)
```

# Shrinkage is adaptive to information

```{r, echo=FALSE}
dd = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
xx = rowSums(dd[,2:5])
dd = dd[xx>0,]
xx = xx[xx>0]
cbind(dd[,1:5],pval,zdat.ash$lfdr)[c(15325,16123),]
```

# Recall FDR problem 1: q values increased by low count genes
```{r,echo=FALSE, fig.height=4,fig.cap=""}
plot(qval.high$qval,qval$qval[xx>1000],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
abline(a=0,b=1,col=2)
```

# ASH q values more robust to inclusion of low count genes

```{r, echo=FALSE,fig.height=3,fig.cap=""}
plot(zdat.ash.high$qvalue,zdat.ash$qvalue[xx>1000],xlab="q values from high only",ylab="q values from all data")
abline(a=0,b=1,col=2,lwd=2)
```

Compare fitted $f(\beta)$, both estimating $\pi_0$ and fixing $\pi_0=0$.
```{r, echo=FALSE,fig.height=4,fig.cap=""}
  x=seq(-4,4,length=100)
  plot(cdf.ash(hh.ash.fdr,x),col=2,type="l",lwd=3,ylab="cdf")
  lines(x,cdf.ash(hh.ash.shrink,x)$y,col=3,lwd=3,lty=2)
```

>>>>

---
title: "test"
output: ioslides_presentation
---
